---
title: "Chr2 SNP Analysis"
author: "Dave Speca"
date: "6/6/2017"
output: html_document
---

```{r}
library(ggplot2)
library(dplyr)
```

###Read in the vcf file and add headers:

```{r, cache=TRUE}
# read in .vcf file
Chr2.data <- read.table("Chr2.test.vcf", as.is=TRUE, na.strings=".")
vcf.header <- system("grep '#C' Chr2.test.vcf",intern = TRUE)
vcf.header <- sub("#","",vcf.header) #get rid of the pound sign
vcf.header <- unlist(strsplit(vcf.header,split="\t"))
colnames(Chr2.data) <- vcf.header
head(Chr2.data)
```

###String split data

```{r}
# Before splitting add NAs to blank cells

Chr2.data$DBA[is.na(Chr2.data$DBA)] <- "NA:NA:NA:NA:NA:NA:NA:NA"

DBA.tmp <- matrix(
  unlist(strsplit(Chr2.data$DBA,split = ":")),
  nrow=nrow(Chr2.data),
  byrow=TRUE
  )
head(DBA.tmp)
colnames(DBA.tmp) <- paste("DBA",c("gt","tot.depth","NObsAllele", "ref.depth","ref.qual","alt.depth","alt.qual","gen.lik"),sep="_")

Chr2.data$Chr2[is.na(Chr2.data$Chr2)] <- "NA:NA:NA:NA:NA:NA:NA:NA"

Chr2.tmp <- matrix(
  unlist(strsplit(Chr2.data$Chr2,split = ":")),
  nrow=nrow(Chr2.data),
  byrow=TRUE
  )
head(Chr2.tmp)
colnames(Chr2.tmp) <- paste("Chr2",c("gt","tot.depth","NObsAllele", "ref.depth","ref.qual","alt.depth","alt.qual","gen.lik"),sep="_")

Chr2.data <- cbind(Chr2.data,DBA.tmp,Chr2.tmp,stringsAsFactors=FALSE)
summary(Chr2.data)
```


###Convert columns back to numeric:
```{r}
Chr2.data[,c("DBA_tot.depth","DBA_ref.depth","DBA_ref.qual","DBA_alt.depth","DBA_alt.qual",
            "Chr2_tot.depth","Chr2_ref.depth","Chr2_ref.qual","Chr2_alt.depth","Chr2_alt.qual")] <- 
  apply(Chr2.data[,c("DBA_tot.depth","DBA_ref.depth","DBA_ref.qual","DBA_alt.depth","DBA_alt.qual",
            "Chr2_tot.depth","Chr2_ref.depth","Chr2_ref.qual","Chr2_alt.depth","Chr2_alt.qual")],
        2,
        as.numeric
        )
head(Chr2.data, 6)
```

###Exploratory analysis:

####Sequencing depth: DBA, raw
```{r}
qplot(Chr2.data$DBA_tot.depth, geom="histogram")
```

####Sequencing depth: DBA, focused
```{r}
qplot(Chr2.data$DBA_tot.depth, geom="histogram", xlim=c(20, 100))
```

####Sequencing depth: Chr2, focused

```{r}
qplot(Chr2.data$Chr2_tot.depth, geom="histogram", xlim=c(20, 100))
```

###Scatterplot of DBA and Chr2 sequencing depths (DP)

```{r}
plot1 <- ggplot(Chr2.data, aes(DBA_tot.depth, Chr2_tot.depth))
plot1 + geom_point() + xlim(20, 600) + ylim(20, 600)
```
###Pretty linear. Good.

###Histogram of QUAL scores:

```{r}
# histogram of low scores
plot2 <- ggplot(Chr2.data, aes(QUAL)) + geom_histogram()
plot2
```
###Well, that's something. Let's try again.

```{r}
# histogram of low scores
plot3 <- ggplot(Chr2.data, aes(QUAL)) + geom_histogram() + xlim(-1, 300)
plot3
```

###Kinda weird that there is a spike at QUAL ~ 60.

###Subset Chr2.data QUAL > 40

```{r}
# subset data keeping only SNPs with quality scores greater than or equal to 40
Chr2.Q40.data <- subset(Chr2.data, QUAL >= 40)
```

###Replot scatterplot of read depths on Q40 dataset:

```{r}
plot4 <- ggplot(Chr2.Q40.data, aes(DBA_tot.depth, Chr2_tot.depth))
plot4 + geom_point() + xlim(20, 600) + ylim(20, 600)
```

###Let's make a table!

```{r}
# count the numbers common and unique to each genotype
ftable(vcf.data.good[,c("IMB211_gt","R500_gt")])
```

